---
title: "Reproducible Research Project 1 - Activity Analysis"
date: "Saturday, August 16, 2014"
output: html_document
---


##Loading and preprocessing the data

```{r, tidy=FALSE}
activityData = read.csv("activity.csv",colClasses=c("integer","Date","integer"))
library(plyr)
activitySum <- ddply(activityData, .(date),colwise(sum))
plot(activitySum$date,activitySum$steps,type="h",xlab="Date",ylab="Total # of Steps",main="Total Steps Taken Per Day Histogram")

```

##Mean total number of steps taken per day

```{r}
activityDailyMean <- ddply(activityData, .(date),colwise(mean),na.rm=T)

activityDailyMean

```


##Median total number of steps taken per day

```{r}

activityDailyMedian <- ddply(activityData, .(date),colwise(median),na.rm=T)

activityDailyMedian

```



##What is the average daily activity pattern?
```{r}
acticiyTimeoftheDayMean <- ddply(activityData, .(interval),colwise(mean),na.rm=T)

acticiyTimeoftheDayMean$interval <- as.POSIXlt(strptime(paste(acticiyTimeoftheDayMean$date,sprintf("%04i",acticiyTimeoftheDayMean$interval),length=4),"%Y-%m-%d%H%M"),tz = "GMT")

plot(acticiyTimeoftheDayMean$interval,acticiyTimeoftheDayMean$steps,t="l",,main="Daily Time of the Day Activity Graph", xlab="Time of the Day", ylab = "Average Steps")

subset(acticiyTimeoftheDayMean,steps == max(acticiyTimeoftheDayMean$steps),select = interval)

```

# The Maximum number of steps are are at interval 8:35 AM


#The number of NA data points are shown below

```{r}
count(activityData[is.na(activityData)])

```

##Imputing missing values

```{r}
count(activityData[is.na(activityData)])

#Fill NAs with mean of steps for that timeoftheday

acticiyTimeoftheDayMeanForFill <- ddply(activityData, .(interval),colwise(mean),na.rm=T)

missVals <- is.na(activityData$steps)

activityDataFill <- activityData

for (i in 1:nrow(activityData))
{
  if (missVals[i])
  {
    activityDataFill[i,1] <- subset(acticiyTimeoftheDayMeanForFill,interval == activityData[i,3])$steps
  }
}

```

## Mean and median is same, before and after imputing, as imputing is based on mean from non.NA observations

```{r}

is.weekday <- function(x = NULL)
{
  if (is.null(x))
    return(NULL)
  
  retVec <- vector(mode = "logical",length = length(x))
  for (i in 1:length(x))
  {
    if (weekdays(x[i]) == "Saturday" | weekdays(x[i]) == "Sunday" )
    {
      retVec[i] <- FALSE
    }
    else
    {
      retVec[i] <- TRUE
    }
  }
  return(retVec)
}

weekdayVec <- is.weekday(activityDataFill$date)

activityDataFillWeekday <- activityDataFill[weekdayVec,]

activityDataFillWeekend <- activityDataFill[!weekdayVec,]


weekdayTimeoftheDayMean <- ddply(activityDataFillWeekday, .(interval),colwise(mean),na.rm=T)

weekdayTimeoftheDayMean$interval <- as.POSIXlt(strptime(paste(weekdayTimeoftheDayMean$date,sprintf("%04i",weekdayTimeoftheDayMean$interval),length=4),"%Y-%m-%d%H%M"),tz = "GMT")



weekendTimeoftheDayMean <- ddply(activityDataFillWeekend, .(interval),colwise(mean),na.rm=T)

weekendTimeoftheDayMean$interval <- as.POSIXlt(strptime(paste(weekendTimeoftheDayMean$date,sprintf("%04i",weekendTimeoftheDayMean$interval),length=4),"%Y-%m-%d%H%M"),tz = "GMT")



par(mfrow=c(2,1))

plot(weekdayTimeoftheDayMean$interval,weekdayTimeoftheDayMean$steps,t="l",main="Weekday Activity Graph", xlab="Time of the Day", ylab = "Average Steps",col="blue")

plot(weekendTimeoftheDayMean$interval,weekendTimeoftheDayMean$steps,t="l",main="Weekend Activity Graph", xlab="Time of the Day", ylab = "Average Steps",col="blue")

```

```{r}
```

```{r}
```

```{r}
```